Paper_Notes

### 2017-10-27 Paper_Notes.Rmd
Summary #4 Commit
Reading Assignment for Monday, October 30

### Atman et al. 2014

@@

>>>>

### Altman et al. 2014

@@

### Bolstad et al. 2003

- Bolstad, Irizarray, Astrand, Speed. "A Comparison of Normalization Methods for High Density Oligonucleotide Array Data Based on Variance and Bias." Bioinformatics 19, no.2 (January 22, 2003): 185-93.

The article compares the performance of three methods of normalization for use when analyzing high density oligonucleotide arrays. It is important to remove sources of variation between arrays of non-biological origin when running experiments that involve multiple high density oligonucleotide arrays. Normalization is a process for reducing unwanted variability. For these arrays, oligonucleotides of 25 base pairs in length are used to probe genes. There are two types of probes: 1) perfect match (PM) probes that match a target sequence exactly and 2) mismatch (MM) probes, which differ from the reference probes only by a single base in the center of the sequence. The normalization methods are performed at the probe intensity level, and are referred to as complete data methods because they make use of data from all arrays in an experiment to form the normalizing relation. These methods are then compared to two methods that make use of a baseline array. The first method is Cyclic loess. It is an approach based upon the idea of the M versus A plot, where M is the difference in log expression values and A is the average of the log expression values. The second method is contrast. The contrast based method is another extension of the M versus A method. Quantile normalization is the last method where the goal is to make the distribution of probe intensities for each array in a set of arrays the same. The results of the study suggest that all three of the complete data methods reduced the variation of the probeset measure across a set of arrays to a greater degree than their reference methods. However, the quantile method performed favorably across all experiments, both in terms of speed and when using their proposed variance and bias criteria.

### Cleveland et al. 1988

- Cleveland, William S, and Susan J Devlin. "Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting." Journal of the American Statistical Association 83, no. 403 (1988): 596-610. - Loess regression. Concept, statistics.

The article discusses locally weighted regression (loess). It is a way of estimating a regression surface through a multivariate smoothing procedure, fitting a function of the independent variables locally and in a moving fashion analogous to how a moving average is computed for a time series. Interestingly, the term loess is derived from the word meaning a deposit of fine clay or silt along river valleys; in a vertical cross-section of Earth, a loess would appear as a narrow, curve-like stratum running through the section. A much wider class of smooth functions can be estimated using local fitting. The applications in the present article illustrate three major uses of the local-fitting methodology. The first is simply to provide an exploratory graphical tool; graphing smooth surfaces that are fitted to the data can give insight into the behavior of the data and help choose parametric models. The second is to provide additional regression diagnostics to check the adequacy of parametric models fitted to the data. The third is to use the loess estimate as the estimated regression surface, without resorting to a parametric class of functions. Other methodologies are discussed, which include multivariate smoothing, the statistical properties of loess that are analogous to those used in the least-squares fitting of parametric functions, several graphical methods that are useful tools for understanding loess estimates and checking assumptions on which the estimation procedure is based, and the M plot.
